---
title: "Lesson 1: Covenants"
---

This will be a lesson on racially restrictive covenants in Roanoke, where students will work with population and spatial data to understand the emergence of systematic residential segregation in the city.

```{r}
#| echo: false

library(sf)
library(tigris)
library(dplyr)
library(ggplot2)
library(leaflet)

options(tigris_use_cache = TRUE)
options(tigris_progress_bar = FALSE)

roanoke <- places(state = "VA", cb = TRUE) %>%
  filter(grepl("Roanoke", NAME))
```

To begin, I made a graph with `ggplot` to show the rapid population growth of Roanoke. I used `annotate` to call attention to the rate of increase. I plan to add racial data as two more lines on this graph. I will use this data to contextualize a 1893 race riot and subsequent efforts to organize the city along racial lines through the use of restrictive covenants.

One thing I obviously need to work out is the the x-axis, which goes up to 2100, even though that is not a year that I have in my data file. Perhaps this is because of the annotation.

```{r}
#| echo: false

roa_pop <- read.csv("roa-pop.csv")

ggplot(roa_pop, aes(x = year, y = population)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(
    title = "Roanoke Population Over Time",
    x = "Year",
    y = "Population"
  ) +
  annotate(
    "label",
    x = max(roa_pop$year, na.rm = TRUE) + 100,
    y = min(roa_pop$population, na.rm = TRUE) + 0.1 * diff(range(roa_pop$population, na.rm = TRUE)),
    label = "Between 1880 and 1890, Roanoke's population grew by more than 2000%",
    fill = "white",
    color = "black",
    label.size = 0.3,
    label.padding = unit(0.3, "lines"),
    hjust = 1,
    vjust = 0,
    size = 4
  ) +
  theme_minimal()
```

For this project, I hope to locate racially restrictive covenants from the Roanoke deeds office. Thinking about how I wanted to map the covenants, I started with `ggplot` and `geo_sf`. I loaded `tigris`, a package that contains the boundaries for Census designated places, and used a few more commands to single out Roanoke.

```{r}
#| echo: false

roanoke <- st_transform(roanoke, 4326)

ggplot(roanoke) +
  geom_sf(fill = "white", color = "black", linewidth = 1) +
  labs(
    title = "Roanoke, VA, City Limits",
  ) +
  theme_void()
```

However, thinking long-term, I decided `ggplot` is perhaps not the best way to visualize spatial data for the tiny parcels that would display where the racially restrictive covenants were. Therefore, I decided to go interactive with `leaflet`, so students can magnify the map. Also, `leaflet` uses OpenStreetMap as a basemap, which would give students several reference points with the existing features on the map.

```{r}
#| echo: false

leaflet(roanoke) %>%
  addTiles() %>%
  addPolygons(
    fillColor = "gray",
    fillOpacity = 0.3,
    color = "black",
    weight = 1
  ) %>%
  setView(lng = -79.9414, lat = 37.2710, zoom = 11)
```

To prepare to map covenants, I figured I would need to download parcel data, which I found from Roanoke's Real Estate Office and have included in my repository as `"parcels.shp"`. Given the massive amount of data in this file, it took quite a long time to run. I therefore decided that I would return to `"parcels.shp"` when I have the data for the covenants. That way, I could isolate the parcels whose deeds had historically contained racially restrictive language, and it would be far less data to load onto `leaflet`.

One issue with my methodology, however, is that I am working with 2022 parcel data and city limits. I plan to work with my Dr. Winling, my advisor, and Max Oickle, who is the GIS specialist for the Chicago Covenants Project, to sort through these questions.

### References

1.  <https://www.w3schools.com/TAGS/tag_iframe.asp>
2.  <https://ggplot2.tidyverse.org>
3.  <https://leafletjs.com>
4.  <https://cran.r-project.org/web/packages/tigris/index.html>

### Lesson Plan

Each lesson contains a PDF document that will house the detailed lesson plan. To do this, I used `<iframe>` to embed my PDF onto the page. This allows users to magnify, download, and print the document. Since PDFs are not accessible, I plan to later display this content with Markdown text. However, I wanted to play around with `<iframe>`.

<iframe src="lesson-template.pdf" width="100%" height="700px">
